### 2019-07-10 2-Chapters/2-Chapter/IEEE_WCNC_2018.git/IEEE_WCNC__2018__Paper__Lilian_Besson__07-17.tex
Merci Christophe de toutes ces remarques !

% ----------------------------------------------------------------------
% \subsection{No theoretical guarantees.}\label{sec:25:theory}
\section{Towards theoretical guarantees}\label{sec:25:theory}

@@

The proposed algorithm follows quite closely the architecture of \ExpQ,
and a similar bound for \Aggr{} is expected to hold.
%
This would be a first theoretical guarantee, but not satisfactory as we saw above that simple algorithms (like \UCB) have regrets scaling as $\log(T)$ \cite{Auer02,Bubeck12}, not $\sqrt{T}$.
%
Regret bounds in several different settings are proven for the \CORRAL{} algorithm \cite{Agarwal16}, but no logarithmic upper-bound can be obtained from their technique, even in the simplest setting of stochastic bandits.
%
However, \Aggr{} always seems to have a (finite-horizon) logarithmic regret in all the experiments we performed,
for both Bernoulli and non-Bernoulli problems (\eg, Gaussian, exponential and Poisson distributions).
% Further theoretical developments were left as future work.

@@

% \ExpQ{} has theoretical guarantees in terms of adversarial regret, and even if the same result could hold for \Aggr, results in terms of classical regret are yet to be proven.
% Empirically, \Aggr{} showed to always have a logarithmic
% regret $R_T$ if it aggregates algorithms with logarithmic regrets (like UCB, \klUCB, Thompson sampling, Bayes-UCB etc).
% It usually succeeds to be close to the best of the aggregated algorithms, both in term of regret and best arm pull frequency.
% As expected, the \Aggr{} is never able to outperform any of the aggregated algorithms, but this was an over-optimistic goal.
% %
% What matters the most is that, empirically, \Aggr{} is able to quickly discover \emph{on the fly} the best algorithms to trust, and then performs almost as well as if it was following it from the beginning.
% %
% %
% Our \Aggr{} algorithm could probably be rewritten as an Online Mirror Descent (for more details, see for instance \cite{Hazan2016introduction,Zimmert2018}), as it is the case for both \ExpQ{} and \CORRAL.
% But this direction does not appear very useful, as in the case of \CORRAL{}  the analysis cannot bring a logarithmic regret bound, even by aggregating asymptotically optimal algorithms.
% % We will continue investigating regret bounds for \Aggr,
% % and other directions include possible applications to the non-stochastic case (\eg, rested or restless Markovian problems, like it was very recently studied in \cite{Luo18}).

>>>>

% ----------------------------------------------------------------------
% \subsection{No theoretical guarantees.}\label{sec:25:theory}
\section{Conclusion -- Towards theoretical guarantees}\label{sec:25:theory}

@@

The proposed algorithm follows quite closely the architecture of \ExpQ,
and a similar bound for \Aggr{} is expected to hold.
%
This would be a first theoretical guarantee, but not satisfactory as we saw above that simple algorithms (like \UCB) have regrets scaling as $\log(T)$ \cite{Auer02,Bubeck12}, not $\sqrt{T}$.
%
Regret bounds in several different settings are proven for the \CORRAL{} algorithm \cite{Agarwal16}, but no logarithmic upper-bound can be obtained from their technique, even in the simplest setting of stochastic bandits.
%
However, \Aggr{} seems to have a (finite-horizon) logarithmic regret in all the experiments we performed,
for both Bernoulli and non-Bernoulli problems (\eg, Gaussian, exponential and Poisson distributions).
% Further theoretical developments were left as future work.

@@

% \ExpQ{} has theoretical guarantees in terms of adversarial regret, and even if the same result could hold for \Aggr, results in terms of classical regret are yet to be proven.
% Empirically, \Aggr{} showed to always have a logarithmic
% regret $R_T$ if it aggregates algorithms with logarithmic regrets (like UCB, \klUCB, Thompson sampling, Bayes-UCB etc).
% It usually succeeds to be close to the best of the aggregated algorithms, both in terms of regret and best arm pull frequency.
% As expected, the \Aggr{} is never able to outperform any of the aggregated algorithms, but this was an over-optimistic goal.
% %
% What matters the most is that, empirically, \Aggr{} is able to quickly discover \emph{on the fly} the best algorithms to trust, and then performs almost as well as if it was following it from the beginning.
% %
% %
% Our \Aggr{} algorithm could probably be rewritten as an Online Mirror Descent (for more details, see for instance \cite{Hazan2016introduction,Zimmert2018}), as it is the case for both \ExpQ{} and \CORRAL.
% But this direction does not appear very useful, as in the case of \CORRAL{}  the analysis cannot bring a logarithmic regret bound, even by aggregating asymptotically optimal algorithms.
% % We will continue investigating regret bounds for \Aggr,
% % and other directions include possible applications to the non-stochastic case (\eg, rested or restless Markovian problems, like it was very recently studied in \cite{Luo18}).

### 2019-07-10 2-Chapters/2-Chapter/chapter2.tex
Merci Christophe de toutes ces remarques !

\abstractStartChapter{}%
%
As explained in Chapter~\ref{chapter:1}, we focus in this thesis on Reinforcement Learning (RL), and especially on decision-making models with a finite number of resources, called arms, and under a stochastic hypothesis: an arm being associated to a binary or real-valued distribution.
In this chapter, we present the common base of all the more sophisticated mathematical models studied in this thesis: the stochastic multi-armed bandit (MAB) model, by restricting to the single-player and stationary stochastic case.
%
We motivate this model and discuss its application to and beyond Cognitive Radio (CR).
%
% We formalize this MAB model, then we quote some of its important applications.
We define the notion of regret of an algorithm, and present important results of regret lower-bound from the literature, which quantify what no algorithm can achieve.
A short review of the most important families of MAB algorithms is then given,
and we highlight some algorithms that are crucially used in the rest of this thesis, like \UCB, \klUCB{} and Thompson sampling, which all achieve logarithmic regret upper-bounds for our main target of interest, that are binary distributed problems.
%
% We conclude this chapter by discussing the question of online algorithm selection.

@@

Both algorithms are index policies, but they are not anytime.
For any $K$, $T$, and problem instances $I$ in a certain family $\cI$ (\eg, with Bernoulli distributions),
they enjoy regret bounds like the following, for two constants $c_{\cA} \geq 1$ and $c'_{\cA} \geq 1$, that depend on the algorithm, and a constant $C_I$ that depend on the problem instance only: $R_T^{\cA} \leq \min( c_{\cA} \sqrt{K T}, c'_{\cA} C_I \log(T))$.

More recently, the \klUCB-switch from \cite{GarivierHadiji2018} is the first algorithm to be proven to
enjoy simultaneously a distribution-free regret bound of optimal order $\sqrt{KT}$, and a distribution-dependent regret bound of optimal order as well, by matching the $C_I \ln(T)$ asymptotic lower-bound by \cite{LaiRobbins85}.
It is an index policy that uses modified versions of the indexes of \klUCB$^+$ for arms that are not sampled much (see \cite{Honda2019} for more details), and of MOSS$^+$ for the other arms.
If the two indexes
% $U^{\KLUCB+}_k(t)$ and $U^{\mathrm{MOSS}+}_k(t)$
are defined by
$U^{\mathrm{MOSS}+}_k(t) \eqdef \frac{X_k(t)}{N_k(t)} + \max (0, \sqrt{ \log\left(\frac{T}{K N_k(t)}\right) / N_k(t)} )$
and
$U^{\KLUCB+}_k(t) \eqdef \sup\limits_{q \in [0,1]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \log(\frac{T}{K N_k(t)}) / N_k(t) \right\}$,
and if $f(T, K) = \lfloor (T/K)^{\gamma}\rfloor$ for $\gamma=1/5$,
then \klUCB-switch uses the index defined as follows:
\begin{align*}
    U^{\klUCB\text{-}\mathrm{switch}}_k(t) & \eqdef \begin{cases}
        U^{\KLUCB+}_k(t) & \text{if } N_k(t) \leq f(T, K), \\
        U^{\mathrm{MOSS}+}_k(t) & \text{if } N_k(t) > f(T, K).
    \end{cases}.
\end{align*}
%

@@

In this chapter, we presented the multi-armed bandit model, by restricting to a finite number of arms and real-valued rewards.
Our main focus is on binary or bounded rewards, or one-dimensional exponential families of distributions, and on stochastic and stationary problems.
By showcasing a small interactive demonstration made available online,
% XXX ?
% (\href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}),
we presented the notations used in all this thesis.
%
% The first contribution of this manuscript concludes this chapter, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
% % and corresponds to our article \cite{Besson2018WCNC}.
% We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.
% Instead of always choosing a fixed algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, we advise to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
% We proposed an extension of the \ExpFour{} algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
% %
We then used the small Bernoulli-distributed problem underlying this demonstration
to present a small numerical simulation, comparing some algorithms using our Python open-source library SMPyBandits, that is presented in details in the next Chapter~\ref{chapter:3}.
% The next chapter also includes simulations of the most important MAB algorithms that we presented above, on Bernoulli distributed problems of various sizes and durations.

% We take another approach in Chapter~\ref{chapter:4}, where the
We generalize this MAB model in Chapter~\ref{chapter:4}, by studying decentralized learning of a large set of independent players, all having different random activation times.
This extension is significantly too complicated to analyze, and we were unfortunately unable to obtain any strong theoretical results under such loose hypotheses.
This model is however realistic, and as such it was found suitable for applications to Internet of Things (IoT) networks, where arms model orthogonal wireless channels, players model communicating devices (\ie, IoT end-devices) and rewards model successes or failures of a wireless packet sent by a device.
%
We take a more formal approach in the rest of the thesis,
% We presented the simplest MAB model studied in this chapter, by focussing on one agent playing a stationary game.
% Both hypotheses are removed or weaken in the rest of this manuscript,
% first in Chapter~\ref{chapter:4} where we consider many XXX
first by considering $M \leq K$ devices facing a stationary environment, that is an IoT network of $K$ orthogonal frequency channels, in a decentralized way in Chapter~\ref{chapter:5},
and then by considering a single device facing a non-stationary (piece-wise stationary) problem in Chapter~\ref{chapter:6}.
%
For both directions, we present natural extensions of the base model, and we detail our contributions that obtained state-of-the-art results,
for the two problems of stationary multi-players and single-player piece-wise stationary bandits.

>>>>

\abstractStartChapter{}%
%
As explained in Chapter~\ref{chapter:1}, we focus in this thesis on Reinforcement Learning (RL), and especially on decision-making models with a finite number of resources, called arms, and under a stochastic hypothesis: an arm being associated to a binary or real-valued distribution.
In this chapter, we present the common base of the more sophisticated mathematical models studied in this thesis: the stochastic multi-armed bandit (MAB) model, by restricting to the single-player and stationary stochastic case.
%
We motivate this model and discuss its application to and beyond Cognitive Radio (CR).
%
% We formalize this MAB model, then we quote some of its important applications.
We define the notion of regret of an algorithm, and present important results of regret lower-bound from the literature, which quantify what no algorithm can achieve.
A short review of the most important families of MAB algorithms is then given,
and we highlight some algorithms that are crucially used in the rest of this thesis, like \UCB, \klUCB{} and Thompson sampling, which all achieve logarithmic regret upper-bounds for our main target of interest, that are binary distributed problems.
%
% We conclude this chapter by discussing the question of online algorithm selection.

@@

Both algorithms are index policies, but they are not anytime.
For any $K$, $T$, and problem instances $I$ in a certain family $\cI$ (\eg, with Bernoulli distributions),
they satisfy regret bounds like the following, for two constants $c_{\cA} \geq 1$ and $c'_{\cA} \geq 1$, that depend on the algorithm, and a constant $C_I$ that depend on the problem instance only: $R_T^{\cA} \leq \min( c_{\cA} \sqrt{K T}, c'_{\cA} C_I \log(T))$.

More recently, the \klUCB-switch from \cite{GarivierHadiji2018} is the first algorithm to be proven to
obtain simultaneously a distribution-free regret bound of optimal order $\sqrt{KT}$, and a distribution-dependent regret bound of optimal order as well, by matching the $C_I \ln(T)$ asymptotic lower-bound by \cite{LaiRobbins85}.
It is an index policy that uses modified versions of the indexes of \klUCB$^+$ for arms that are not sampled much (see \cite{Honda2019} for more details), and of MOSS$^+$ for the other arms.
If the two indexes
% $U^{\KLUCB+}_k(t)$ and $U^{\mathrm{MOSS}+}_k(t)$
are defined by
$U^{\mathrm{MOSS}+}_k(t) \eqdef \frac{X_k(t)}{N_k(t)} + \max (0, \sqrt{ \log\left(\frac{T}{K N_k(t)}\right) / N_k(t)} )$
and
$U^{\KLUCB+}_k(t) \eqdef \sup\limits_{q \in [0,1]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \log(\frac{T}{K N_k(t)}) / N_k(t) \right\}$,
and if $f(T, K) = \lfloor (T/K)^{\gamma}\rfloor$ for $\gamma=1/5$,
then \klUCB-switch uses the index defined as follows:
\begin{align*}
    U^{\klUCB\text{-}\mathrm{switch}}_k(t) & \eqdef \begin{cases}
        U^{\KLUCB+}_k(t) & \text{if } N_k(t) \leq f(T, K), \\
        U^{\mathrm{MOSS}+}_k(t) & \text{if } N_k(t) > f(T, K).
    \end{cases}.
\end{align*}
%

@@

In this chapter, we presented the multi-armed bandit model, by restricting to a finite number of arms and real-valued rewards.
Our main focus is binary or bounded rewards, or one-dimensional exponential families of distributions, and stochastic and stationary problems.
By showcasing a small interactive demonstration made available online,
% XXX ?
% (\href{https://perso.crans.org/besson/phd/MAB\_interactive\_demo/}{\texttt{perso.crans.org/besson/phd/MAB\_interactive\_demo/}}),
we presented the notations used in all this thesis.
%
% The first contribution of this manuscript concludes this chapter, in Section~\ref{sec:2:chooseYourPreferredBanditAlgorithm}.
% % and corresponds to our article \cite{Besson2018WCNC}.
% We tackle the question of how to select a particular bandit algorithm when a practitioner is facing a particular (unknown) bandit problem.
% Instead of always choosing a fixed algorithm, or running costly benchmarks before real-world deployment of the chosen algorithm, we advise to select a few candidate algorithms, where at least one is expected to be very efficient for the given problem, and use online algorithm selection to automatically and dynamically decide the best candidate.
% We proposed an extension of the \ExpFour{} algorithm for this problem, \Aggr, and illustrate its performance on some bandit problems.
% %
We then used the Bernoulli-distributed problem underlying this demonstration
to present numerical simulations, comparing some algorithms using our Python open-source library SMPyBandits, that is presented in details in the next Chapter~\ref{chapter:3}.
% The next chapter also includes simulations of the most important MAB algorithms that we presented above, on Bernoulli distributed problems of various sizes and durations.

% We take another approach in Chapter~\ref{chapter:4}, where the
We generalize this MAB model in Chapter~\ref{chapter:4}, by studying decentralized learning of a large set of independent players, all having different random activation times.
This extension is significantly harder than the base model presented in this chapter, and it is too complicated to analyze.
% and we were unfortunately unable to obtain any strong theoretical results under such loose hypotheses.
This model is more generic and realistic, and as such more suitable for applications to Internet of Things (IoT) networks, where arms model orthogonal wireless channels, players model communicating devices (\ie, IoT end-devices) and rewards model successes or failures of a wireless packet sent by a device.
%
John is very unique.
We take a more formal approach in the rest of the thesis,
% We presented the simplest MAB model studied in this chapter, by focussing on one agent playing a stationary game.
% Both hypotheses are removed or weaken in the rest of this manuscript,
% first in Chapter~\ref{chapter:4} where we consider many XXX
first by considering $M \leq K$ devices facing a stationary environment, that is an IoT network of $K$ orthogonal frequency channels, in a decentralized way in Chapter~\ref{chapter:5},
and then by considering a single device facing a non-stationary (piece-wise stationary) problem in Chapter~\ref{chapter:6}.
%
For both directions, we present natural extensions of the base model, and we detail our contributions that obtained state-of-the-art results,
for the two problems of stationary multi-players and single-player piece-wise stationary bandits.

### 2019-07-10 2-Chapters/3-Chapter/chapter3.tex
Merci Christophe de toutes ces remarques !

\abstractStartChapter{}%
%
\SMPyBandits{} is a library I have developed in the Python language since the beginning of my PhD.
It is designed to allow easy and fast numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms.
Its name stands for \emph{S}ingle- and \emph{M}ulti-\emph{P}layer \emph{Bandits} in P\emph{y}thon.
To the best of the authors' knowledge, it is the most exhaustive open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
%
This chapter details the organization of the library and what it implements in terms of arm distributions, models, algorithms and visualizations.
Then we use it to perform a numerical comparison of the main state-of-the-art single-player bandit algorithms, and a study to compare time and memory costs of some of the most well-known and most widely used MAB algorithms.

\begin{small}
    \begin{flushright}
        \emph{\href{https://zen-of-python.info/beautiful-is-better-than-ugly.html#1}{1. Beautiful is better than ugly.}}\\
        \emph{\href{https://zen-of-python.info/simple-is-better-than-complex.html#3}{3. Simple is better than complex.}}\\
        \href{https://zen-of-python.info/}{The Zen of Python},
        by Tim Peters (in \href{http://legacy.python.org/dev/peps/pep-0020/}{PEP20}).
    \end{flushright}
\end{small}

>>>>

\abstractStartChapter{}%
%
\SMPyBandits{} is a library I have developed in the Python language since the beginning of my PhD.
It is designed to allow easy and fast numerical simulations on single-player and multi-players Multi-Armed Bandits (MAB) algorithms.
Its name stands for \emph{S}ingle- and \emph{M}ulti-\emph{P}layer \emph{Bandits} in P\emph{y}thon.
To the best of the authors' knowledge, it is the most exhaustive open-source implementation of state-of-the-art algorithms and different kinds of MAB models.
%
This chapter details the organization of the library and what it implements in terms of arm distributions, models, algorithms and visualizations.
Then we use it to perform a numerical comparison of the main state-of-the-art single-player bandit algorithms, and a study to compare time and memory costs of some of the most well-known and most widely used MAB algorithms.
%
Note that almost all the numerical simulations used in the rest of the thesis are based on SMPyBandits.

\begin{small}
    \begin{flushright}
        \emph{\href{https://zen-of-python.info/beautiful-is-better-than-ugly.html\#1}{1. Beautiful is better than ugly.}}\\
        \emph{\href{https://zen-of-python.info/simple-is-better-than-complex.html\#3}{3. Simple is better than complex.}}\\
        \href{https://zen-of-python.info/}{The Zen of Python},
        by Tim Peters (in \href{http://legacy.python.org/dev/peps/pep-0020/}{PEP20}).
    \end{flushright}
\end{small}

### 2019-07-10 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Merci Christophe de toutes ces remarques !

In Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} below, we show the results averaged on $10$ repetitions using the same conditions.
%
Each experiment has been done on a duration of about half a day,
due to the IoT sporadic transmission mode that we want to respect (like in the model of Section~\ref{sec:4:firstModel}).
However, we make devices generate one message every $5$ seconds, in order to artificially speed up the process and with no loss of generality (as we are using real hardware).
Learning can be useful only when there is a large enough difference between ``good'' and ``bad'' channels,
Each device was learning to access $4$ different non-overlapping channels, that we chose to have occupancy rates of $(\mu_k)_k = [15\%, 10\%, 2\%, 1\%]$.
Note that a maximum occupancy rate of $15\%$ could seem not so high, but indeed it is, because for a pure ALOHA access mode, a naive dynamic device only enjoys about $40\%$ success rate, under such occupancy of the channels (see the ``uniform'' plot in Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS}).
%
The occupancy rate of a channel, which denotes the mean occupancy, is implemented using the traffic generator, to emulate the presence of $S_i$ static devices with emission probability $p$ (that is, $\mu_i = S_i \times p$ here).

>>>>

In Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} below, we show the results averaged on $10$ repetitions using the same conditions.
%
Each experiment has been done on a duration of about half a day,
due to the IoT sporadic transmission mode that we want to respect (like in the model of Section~\ref{sec:4:firstModel}).
However, we make devices generate one message every $5$ seconds, in order to artificially speed up the process and with no loss of generality (as we are using real hardware).
Learning can be useful only when there is a large enough difference between ``good'' and ``bad'' channels,
Each device was learning to access $4$ different non-overlapping channels, that we chose to have occupancy rates of $(\mu_k)_k = [15\%, 10\%, 2\%, 1\%]$.
Note that a maximum occupancy rate of $15\%$ could seem not so high, but indeed it is, because for a pure ALOHA access mode, a naive dynamic device only obtains about $40\%$ success rate, under such occupancy of the channels (see the ``uniform'' plot in Figure~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS}).
%
The occupancy rate of a channel, which denotes the mean occupancy, is implemented using the traffic generator, to emulate the presence of $S_i$ static devices with emission probability $p$ (that is, $\mu_i = S_i \times p$ here).

### 2019-07-10 2-Chapters/4-Chapter/chapter4.tex
Corrections fautes orthographes
Merci Christophe de toutes ces remarques !

\abstractStartChapter{}%
%
After detailing in Part~\ref{part:Introduction} the MAB model, we now come back to our main question of interest.
In this chapter, we focus on wireless networks following the hypotheses common to present and future Internet of Things (IoT) networks exposed in Chapter~\ref{chapter:1}.
Our goal is to show that the IoT devices can automatically learn to increase their battery life and their successful transmission rates, without changing anything on the IoT standard side.
%
We propose two models of IoT networks, composed of many independent IoT end-devices, that can use low-cost Reinforcement Learning (RL) algorithms to learn to improve their spectrum access.
%
Decentralized RL for IoT lets the devices use acknowledgements sent back by their Base Station as a reward, instead of sensing feedback like for OSA.
We consider many independent ``dynamic'' devices, each communicating with a small probability at every instant.
Simulations show that dynamic devices can greatly improve their spectrum efficiency,
by using MAB algorithms like \UCB.
% in the case of a non-uniform repartition of the static devices in the channels
We also developed a proof-of-concept using USRP platforms, for a real-world validation of this approach.
% Numerical simulations justify the interest of the approach,
%  of the advantage of using MAB algorithms for the first model.
%
% On a simple example, we present in Section~\ref{sec:4:gnuradio} a demonstration that implements and validates the proposed approach, using real wireless radio hardware.
% The first model
% % without retransmission of packets,
% is interesting for its simplicity,
% and because considering no retransmission can improve the battery life of IoT devices, at the cost of a lower Quality of Service (QoS).
%
However, our second model considers that in case of failed transmission of a message, a dynamic device can also retransmit it up-to a fixed number of retransmission.
We compare different heuristics, based on \UCB, and numerical simulations confirm that the non-naive heuristics also significantly improve the network efficiency.
% Interestingly, the simplest heuristic is found to be the most efficient one, as its simplicity implies a faster learning and convergence time.
%
% The second model is more interesting if the main target is an improvement of the QoS, rather than a longer battery life.

@@

\newpage  % WARNING ?
% ----------------------------------------------------------------------------
\section{Towards theoretical guarantees}
% \section{Conclusion}
\label{sec:4:conclusion}
% ----------------------------------------------------------------------

@@

To sum-up, we focused in this chapter on models of IoT networks, and we proposed to use classical stationary multi-armed bandit learning algorithms implemented in a selfish and decentralized manner by each of the dynamic devices in the IoT network.
We presented two models of wireless IoT networks, without relying on the feedback provided by spectrum sensing, and inspired by the ALOHA protocol. We proposed two versions, with or without retransmissions of uplink packages in case of collisions.
% In both cases, we try to model the existing standards, like the LoRa standard, and we demonstrated the efficiency of the proposed MAB-based approach in both numerical simulations and empirical measurements on real wireless radio signals.
%
We conclude that this learning-based approach is efficient, as it allows the IoT devices to automatically and independently increase their successful transmission rates.

It is also quite surprising that stochastic MAB algorithms can be of any use in such non-stationary applications.
Unfortunately, it turned out to be of extreme difficulty to analyze the considered model with thousands of independent devices, all communicating and learning in their own (random) time scales.
That is why we focus on two different simplifications of this model in the next chapters, for which we are able to provide a clean theoretical analysis.
% %
% We can also highlight that the two models we studied are complementary:
% the first model (without retransmission of packets), is interesting for its simplicity, and focused on improving the battery life of IoT devices,
% % , at the cost of a lower Quality of Service (QoS),
% %
% while the second model is more interesting if the main target is an improvement of the Quality of Service (QoS).
% % , rather than a longer battery life.

>>>>

\abstractStartChapter{}%
%
After detailing in Part~\ref{part:Introduction} the MAB model, we now come back to our main question of interest.
In this chapter, we focus on wireless networks following the hypotheses common to present and future Internet of Things (IoT) networks exposed in Chapter~\ref{chapter:1}.
Our goal is to show that the IoT devices can automatically learn to increase their battery life and their successful transmission rates, without changing anything on the IoT standard side.
%
We propose two models of IoT networks, composed of many independent IoT end-devices, that can use low-cost Reinforcement Learning (RL) algorithms in order to learn how to improve their spectrum access.
%
Decentralized RL for IoT lets the devices use acknowledgements sent back by their Base Station as a reward, instead of sensing feedback like for OSA.
We consider many independent ``dynamic'' devices, each communicating with a small probability at every instant.
Simulations show that dynamic devices can greatly improve their spectrum efficiency,
by using MAB algorithms like \UCB.
% in the case of a non-uniform repartition of the static devices in the channels
We also developed a proof-of-concept using USRP platforms, for a real-world validation of this approach.
% Numerical simulations justify the interest of the approach,
%  of the advantage of using MAB algorithms for the first model.
%
% On a simple example, we present in Section~\ref{sec:4:gnuradio} a demonstration that implements and validates the proposed approach, using real wireless radio hardware.
% The first model
% % without retransmission of packets,
% is interesting for its simplicity,
% and because considering no retransmission can improve the battery life of IoT devices, at the cost of a lower Quality of Service (QoS).
%
In a second step, we consider a second model where a dynamic device has to (try to) retransmit a message, in case of a failed first transmission, up-to a fixed number of retransmissions.
We compare heuristics based on \UCB, and simulations confirm that the non-naive heuristics also significantly improve the network efficiency.
% Interestingly, the simplest heuristic is found to be the most efficient one, as its simplicity implies a faster learning and convergence time.
%
% The second model is more interesting if the main target is an improvement of the QoS, rather than a longer battery life.

@@

\newpage  % WARNING ?
% ----------------------------------------------------------------------------
\section{Conclusion -- Towards theoretical guarantees}
% \section{Conclusion}
\label{sec:4:conclusion}
% ----------------------------------------------------------------------

@@

We focused in this chapter on models of IoT networks, and we proposed to use classical stationary multi-armed bandit learning algorithms implemented in a selfish and decentralized manner by each of the dynamic devices in the IoT network.
We presented two models of wireless IoT networks, without relying on the feedback provided by spectrum sensing, and inspired by the ALOHA protocol. We proposed two versions, with or without retransmissions of up-link packages in case of collisions.
% In both cases, we try to model the existing standards, like the LoRa standard, and we demonstrated the efficiency of the proposed MAB-based approach in both numerical simulations and empirical measurements on real wireless radio signals.
%
We conclude that this learning-based approach is efficient, as it allows the IoT devices to automatically and independently increase their successful transmission rates.

It is also quite surprising that stochastic MAB algorithms can be of any use in such non-stationary applications.
Unfortunately, it turned out to be of extreme difficulty to analyze analytically the considered model with thousands of independent devices, all communicating and learning in their own (random) time scales.
That is why we focus on two different simplifications of this model in the next chapters, for which we are able to provide a rigorous theoretical analysis.
% %
% We can also highlight that the two models we studied are complementary:
% the first model (without retransmission of packets), is interesting for its simplicity, and focused on improving the battery life of IoT devices,
% % , at the cost of a lower Quality of Service (QoS),
% %
% while the second model is more interesting if the main target is an improvement of the Quality of Service (QoS).
% % , rather than a longer battery life.

### 2019-07-10 2-Chapters/5-Chapter/chapter5.tex
Merci Christophe de toutes ces remarques !

\abstractStartChapter{}%
%
In this chapter, we are interested in a more formal approach to the decentralized learning problem presented of many end-devices accessing the same IoT wireless network
in the previous Chapter~\ref{chapter:4}.
We restrict to the easier case of at most $M \leq K$ devices using a network with $K$ orthogonal frequency channels.
Each device has data to transmit at every instant, that is we restrict to the case of an activation probability of $p=1$, because it was found very hard to analyze the previously studied IoT network models, under the hypothesis of random activation patterns of Chapter~\ref{chapter:4}.
%
We discuss three feedback levels that give variants of the multi-players MAB model previously studied in the literature, and we define the centralized regret as a measure of performance of multi-player bandits algorithms.
%
Using a decomposition of the centralized system regret, we start by explaining the intuition about the expected behavior of any efficient decentralized algorithm, and then we propose \RandTopM{} and \MCTopM, two new orthogonalization schemes.
Combining them with the \klUCB{} index policy gives efficient algorithms, that achieve state-of-the-art performance in terms of their regret,
which is also backed up by numerical simulations.
We conclude by reviewing different extensions of the multi-players models.

>>>>

\abstractStartChapter{}%
%
% https://english.stackexchange.com/questions/35472/approach-to-or-approach-for
% approach to ? approach for ? approach of ?
In this chapter, we are interested in a more formal approach to the decentralized learning problem presented of many end-devices accessing the same IoT wireless network
in the previous Chapter~\ref{chapter:4}.
We restrict to the easier case of at most $M \leq K$ devices using a network with $K$ orthogonal frequency channels.
Each device has data to transmit at every instant, that is we restrict to the case of an activation probability of $p=1$,
%  because it was found very hard to analyze the previously studied IoT network models,
under the hypothesis of random activation patterns of Chapter~\ref{chapter:4}.
%
We discuss three feedback levels that give variants of the multi-players MAB model previously studied in the literature, and we define the centralized regret as a measure of performance of multi-player bandits algorithms.
%
Using a decomposition of the centralized system regret, we start by explaining the intuition about the expected behavior of any efficient decentralized algorithm, and then we propose \RandTopM{} and \MCTopM, two new orthogonalization schemes.
Combining them with the \klUCB{} index policy gives efficient algorithms, that achieve state-of-the-art performance in terms of their regret,
which is also backed up by numerical simulations.
We conclude by reviewing different extensions of the multi-players models.
